Journal and article coverage

### 2025-03-05 analysis/manuscript.qmd
Finish first draft of coverage analysis, next replication

jn_upset_publisher_ind <- 
  jn_upset_publisher |> 
  filter(Crossref == TRUE, Scopus == TRUE, WoS == TRUE) |> 
  group_by(publisher) |> 
  summarise(n_journals = n_distinct(issn_l), articles = sum(overlap))
```

Figure \ref{fig-upset_coverage_results_publisher} illustrates the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry between 2019 and 2023.
Elsevier was best represented in the union set of all investigated databases, with `r jn_upset_publisher_ind |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals with at least one open access article that published `r jn_upset_publisher_ind |> filter(publisher == "Elsevier") |> pull(articles)` articles between the five-year period. Springer Nature followed with `r jn_upset_publisher_ind |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals publishing `r jn_upset_publisher_ind |> filter(publisher == "Springer Nature") |> pull(articles)` articles, and Wiley with `r jn_upset_publisher_ind |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals and `r jn_upset_publisher_ind |> filter(publisher == "Wiley") |> pull(articles)` articles.
The remaining `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(esac_publisher) |> nrow()` publishers recorded  a total of `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(issn_l) |> nrow()` hybrid journals that published  `r jn_upset_publisher_ind |> filter(publisher == "Other") |> pull(articles)` articles.

Notably, the three largest publishers--Elsevier, Springer Nature and Wiley--were well represented in the exclusive intersection of all three data sources examined.
In terms of journal coverage, they accounted for `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`%) of the hybrid journals included in the intersection.
This dominance is even more pronounced in terms of article coverage, with these publishers having a joint indexing coverage of `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()` articles (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as measured by shared DOIs.

@@

When considering journals exlcusively covered by Scopus and Web of Science

When considering journals exlcusively covered by hoaddata

>>>>

jn_upset_publisher_ind <- 
  jn_upset_publisher |>
  filter(Crossref == TRUE, Scopus == TRUE, WoS == TRUE) |> 
  group_by(publisher) |>
  summarise(n_journals = n_distinct(issn_l), articles = sum(overlap, na.rm = TRUE))

jn_upset_publisher_all <- 
  jn_upset_publisher |>
  group_by(publisher) |>
  summarise(n_journals = n_distinct(issn_l), articles = sum(overlap, na.rm = TRUE))

jn_upset_publisher_all <- jn_upset_publisher |>
  summarise(n_journals_all = n_distinct(issn_l), articles_all = sum(overlap, na.rm = TRUE)) |>
  bind_cols(jn_upset_publisher_all) |>
  mutate(prop_jn = n_journals / n_journals_all,
         prop_art = articles / articles_all)
```

Figure \ref{fig-upset_coverage_results_publisher} presents the coverage of hybrid journals in transformative agreements across data sources from 2019 to 2023 with a focus on publisher portfolios.
The analysis highlights the dominance of the three largest publishers--Elsevier, Springer Nature, and Wiley--, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of articles published during this five-year period.
In terms of article volume, Elsevier led with `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(articles)` artricles (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_art)*100`% of the total) published across `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals(`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_jn)*100`% of the total). 
Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%), while Wiley recorded `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
The remaining `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(esac_publisher) |> nrow()` publishers collectively accounted for `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`%).

@@

The three largest publishers--Elsevier, Springer Nature and Wiley--were best represented in the exclusive intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage with `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()` (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as determined through shared DOIs.
When examining publication volume per journal, (\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals, followed by Springer Nature and Wiley.

A comparative analysis of publisher portfolios across different indexing sets demonstrates that publisher were not represented uniformly. 
Springer Nature exhibited `r jn_upset_publisher |> filter(esac_publisher == "Springer Nature", WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow()` hybrid journals exclusively indexed in both hoaddata and Scopus.
This set included journals from the Chinese Academy of Science, German-language medical journals, and Eastern European publications including the *Journal of Mathematical Sciences*, which also publishes English-language translations of Russian-language works. 
Additionally, this subset included titles with a more broader disciplianry focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
The set also captured ceased journals, providing insights into the dynamic nature of journal portfolios and business strategies.

```{r miss_portfolios}
miss_tas <- jn_upset_publisher |> 
  filter(Crossref == FALSE) |> 
  group_by(esac_publisher) |> 
  summarise(journals = n_distinct(issn_l),
            n = sum(overlap, na.rm = TRUE)) |> 
  arrange(desc(n))

cr_tas <- jn_upset_publisher |>
  filter(Crossref == TRUE) |> 
  pull(esac_publisher)

miss_publishers <- miss_tas |>
  filter(!esac_publisher %in% cr_tas)
```

Further examination of publisher portfolios not covered by hoaddata but present in Scopus or Web of Science identified several publishers with missing Creative Commons (CC) license information in Crossref.
In particular, `r miss_tas$esac_publisher[1]` represented `r miss_tas$journals[1]` journals with `r miss_tas$n[1]` articles, `r miss_tas$esac_publisher[2]` accounted for `r miss_tas$journals[2]` journals with `r miss_tas$n[2]` articles, and `r miss_tas$esac_publisher[3]` recorded `r miss_tas$journals[3]` journals and `r miss_tas$n[3]` articles.
In total, `r nrow(miss_publishers)` publishers were not represented in Crossref.

An inspection of individual journals uncovered discrepancies in Unpaywall's open access identification for certain publishers that typically deposit CC licenses with Crossref.
Notably, some subscription-only journals contained one or two articles erroneously tagged as hybrid open access by Unpaywall, which were subsequently reflected in Scopus and Web of Science. Examples of such misclassifications include Elsevier's *Journal of Bioscience and Bioengineering* and Springer Nature's *Journal of Mechanical Science and Technology*.

### 2025-03-05 analysis/manuscript.tex
Finish first draft of coverage analysis, next replication

Figure \ref{fig-upset_coverage_results_publisher} illustrates the
coverage of hybrid journals in transformative agreements by publisher
portfolio according to the ESAC registry between 2019 and 2023. Elsevier
was best represented in the union set of all investigated databases,
with 1,672 hybrid journals with at least one open access article that
published 2,345,205 articles between the five-year period. Springer
Nature followed with 1,561 hybrid journals publishing 1,058,564
articles, and Wiley with 1,151 hybrid journals and 770,546 articles. The
remaining 54 publishers recorded a total of 6,476 hybrid journals that
published 2,115,372 articles.

Notably, the three largest publishers--Elsevier, Springer Nature and
Wiley--were well represented in the exclusive intersection of all three
data sources examined. In terms of journal coverage, they accounted for
4,384 (55\%) of the hybrid journals included in the intersection. This
dominance is even more pronounced in terms of article coverage, with
these publishers having a joint indexing coverage of 4,174,315 articles
(66\%).

@@

When considering journals exlcusively covered by Scopus and Web of
Science

When considering journals exlcusively covered by hoaddata

>>>>

Figure \ref{fig-upset_coverage_results_publisher} presents the coverage
of hybrid journals in transformative agreements across data sources from
2019 to 2023 with a focus on publisher portfolios. The analysis
highlights the dominance of the three largest publishers--Elsevier,
Springer Nature, and Wiley--, which collectively accounted for 47\% of
hybrid journals and 62\% of articles published during this five-year
period. In terms of article volume, Elsevier led with 2,441,358
artricles (33\% of the total) published across 1,951 hybrid
journals(16\% of the total). Springer Nature followed with 1,247,578
articles (17\%) in 2,311 hybrid journals (19\%), while Wiley recorded
858,939 articles (12\%) in1,382 hybrid journals (11\%). The remaining 54
publishers collectively accounted for 2,749,847 articles (38\%) in 6,476
hybrid journals (53\%).

@@

The three largest publishers--Elsevier, Springer Nature and Wiley--were
best represented in the exclusive intersection of all three data sources
(hoaddata, Scopus, and Web of Science). Together, they comprised 4,384
hybrid journals (55\% of the intersectional set) and dominated article
coverage with 4,174,315 (66\%), as determined through shared DOIs. When
examining publication volume per journal,
(\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
average, the largest journals, followed by Springer Nature and Wiley.

A comparative analysis of publisher portfolios across different indexing
sets demonstrates that publisher were not represented uniformly.
Springer Nature exhibited 519 hybrid journals exclusively indexed in
both hoaddata and Scopus. This set included journals from the Chinese
Academy of Science, German-language medical journals, and Eastern
European publications including the \emph{Journal of Mathematical
Sciences}, which also publishes English-language translations of
Russian-language works. Additionally, this subset included titles with a
more broader disciplianry focus such as \emph{SN Computer Science} and
newly launched hybrid journals like \emph{Nature Computational Science},
which started in 2021 and was indexed in Scopus but not yet in Web of
Science. The set also captured ceased journals, providing insights into
the dynamic nature of journal portfolios and business strategies.

Further examination of publisher portfolios not covered by hoaddata but
present in Scopus or Web of Science identified several publishers with
missing Creative Commons (CC) license information in Crossref. In
particular, Emerald represented 322 journals with 86,409 articles, AIP
Publishing accounted for 24 journals with 64,898 articles, and World
Scientific recorded 87 journals and 42,531 articles. In total, 9
publishers were not represented in Crossref.

An inspection of individual journals uncovered discrepancies in
Unpaywall's open access identification for certain publishers that
typically deposit CC licenses with Crossref. Notably, some
subscription-only journals contained one or two articles erroneously
tagged as hybrid open access by Unpaywall, which were subsequently
reflected in Scopus and Web of Science. Examples of such
misclassifications include Elsevier's \emph{Journal of Bioscience and
Bioengineering} and Springer Nature's \emph{Journal of Mechanical
Science and Technology}.
